---
title: "Assignment 2 Question 4"
output: pdf_document
---

# a).
```{r}
# generate two predictor variables
x1 <- rnorm(200, mean=0, sd=2)
x2 <- rnorm(200, mean=0, sd=2)

# generate epsilon
eps <- rnorm(200, mean=0, sd=1)

# generate response vector y
y = 1 + 2*x1 + 5*x2 + eps

# estimate the least squares regression line
reg <- lm(y ~ x1+x2)
summary(reg)
```

**the estimated line is:**
```{r}
paste("y=",
      reg$coefficients[1],"+",
      reg$coefficients[2],"x1", "+",
      reg$coefficients[3],"x2")
```

\newpage
# b).


```{r}
# estimate the variance for each of the three regression coefficients
reg_sum <-summary(reg)
(reg_sum$coefficients[1,2])^2
(reg_sum$coefficients[2,2])^2
(reg_sum$coefficients[3,2])^2
```


```{r}
# theoretical values for the variance of these predictors
X <- cbind(rep(1, 200), x1, x2)
diag(solve(t(X) %*% X))
```

**the estimated variance of the three regression coefficients are close to**
**theoretical values for the variance of these predictors.**

\newpage
# c).
```{r}
paste("p-value is",
      reg_sum$coefficients[2,4])
```
**p-value is less than significant level 0.05, so we reject null hypothesis.**
**we can conclude that the coefficient for x1 is not zero.**

\newpage
# d).
```{r}
# repeat part a. and c. 1000 times
count <- 0
for(i in 1:1000){
  x1 <- rnorm(200, mean=0, sd=2)
  x2 <- rnorm(200, mean=0, sd=2)
  eps <- rnorm(200, mean=0, sd=1)
  y = 1 + 2*x1 + 5*x2 + eps
  reg <- lm(y ~ x1+x2)
  reg_sum <- summary(reg)
  pval <- reg_sum$coefficients[2,4]
  if (pval < 0.05) count <- count + 1
}
print(count)
```
**repeat part a and c 1000 times, we reject null hypothesis 1000 times. **








